### 2022-12-06 paper.rmd

---
title: "IML Term project paper"
author: "Julia Palorinne, Pyry Silomaa, Sara Sippola"
output: pdf_document
urlcolor: blue
date: '`r format(Sys.Date(), "%d.%m.%Y")`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(kableExtra)
```

```{r load data, echo=FALSE}
npfOrig <- read.csv("npf_train.csv")
# Create class2
npfOrig$class2[npfOrig$class4 == "nonevent"] <- 0
npfOrig$class2[npfOrig$class4 != "nonevent"] <- 1
# Add date as row name
rownames(npfOrig) <- npfOrig[, "date"]
# Remove columns id, date, class4, partlybad
npfData <- npfOrig[, -(1:4)]
```

# Introduction

In this project we trained a classifier on a data set of atmospheric measurements. The task is to predict whether new particle formation (NPF) happens or not on a given day based on the atmospheric data.

# Preprocessing of the data

## Initial data analysis

The training data consists of several variables measured on 464 non-consecutive days. The variables are daily means and standard deviations of measurements such as carbon dioxide concentration, solar radiation and air temperature. Some of the variables are of the same phenomenon measured at different heights.

Many classifiers are affected by correlation and colinearity between variables. As expected, we found that many of the variables describing the same phenomenon are correlated, as are variables related to radiation (see figure below). We take a more detailed look into this in the section *Correlation between parameters*.

We also familiarized ourselves with the data through the [smear.avaa.scs.fi](https://smear.avaa.csc.fi/) webpage that offers further visualizations details about the data and the measurement site, and with the variable names and details available at [wiki.helsinki.fi](https://wiki.helsinki.fi/pages/viewpage.action?pageId=243959901).

(The data is over non-consecutive days, ok, but are there any significant gaps e.g. a month?)

```{r original correlation plot, echo=FALSE, fig.dim=c(7, 7), message=FALSE}
library(corrplot)
# Original correlations
cmData <- cor(npfData)
corrplot(cmData, order = "FPC", tl.cex = 0.3, tl.col = "black")
```

## Correlation between parameters

The initial data analysis showed that many of the parameters are correlated. For many classifiers, highly correlated parameters are problematic, and for this reason we set out to drop some of the most highly correlated columns from the data set. We know that many of the parameters are measurements of the same thing at different heights. For example, water vapour concentration (H20) has been measured at heights 4.2m, 8.4m, 16.8m, 33.6m, 50.4m and 67.2m. The correlation coefficients between the daily means at different heights are essentially 1 (see table 1).

Hence we decided to discard most of the measurements of H2O and other variables for which we had multiple correlating measurements, both the means and the standard deviations. We ended up keeping the measurements measured at 16.8 meters as for some parameters, the measurements were taken only at that height. We found that after this, there are 38 unique columns. The figure below describes the correlations after omitting all but the measurements taken at the selected height.

```{r H20 correlations, echo=FALSE, fig.dim=c(7, 7), message=FALSE}
kable(cmData[startsWith(colnames(cmData), "H2O") & endsWith(colnames(cmData), ".mean"), 
             startsWith(colnames(cmData), "H2O")  & endsWith(colnames(cmData), ".mean")],
      caption = "Correlation (H20)") %>%
  kable_styling(latex_options = "HOLD_position")
```

```{r correlation plot 168, echo=FALSE, fig.dim=c(6, 6)}
# Select global and 16.8 m high values for correlation plot
npf168 <- data.frame(npfOrig[, (5:6)], npfOrig[, (13:16)], npfOrig[, (27:30)], 
                  npfOrig[, (41:42)], npfOrig[, (53:54)], npfOrig[, (63:72)],
                  npfOrig[, (83:88)], npfOrig[, (99:105)])
cm168 <- cor(npf168)
corrplot(cm168, order = "FPC", tl.cex = 0.5, tl.col = "black")
```

The correlation plot for the remaining parameter shows us that the remaining hihgly correlated parameters are all radiation-related.

# Classifier

## Description of considered machine learning approaches

```{r preprocess data, echo=FALSE}
# Sample half of the data for training and testing sets
set.seed(42)
idx <- sample(nrow(npfData), nrow(npfData)/2)
npf_train <- npfData[idx, ]
npf_test <- npfData[-idx, ]
```

## Chosen calssifier, pros and cons of this particular classifier for this application

# Results

## Classifier performance (numerical)

## Insights, conclusions, discussion etc.