### 2022-12-11 final.rmd
trees

---
title: "IML Term project"
author: "Julia Palorinne, Pyry Silomaa, Sara Sippola"
output: pdf_document
date: '`r format(Sys.Date(), "%d.%m.%Y")`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(kableExtra)
```

```{r load data}
npfOrig <- read.csv("npf_train.csv")
npfOrig$class2 <- ifelse(npfOrig$class4 == "nonevent", 0, 1)
rownames(npfOrig) <- npfOrig[, "date"]
npf.168 <- data.frame(npfOrig[, (5:6)], npfOrig[, (13:16)], npfOrig[, (27:30)], 
                  npfOrig[, (41:42)], npfOrig[, (53:54)], npfOrig[, (63:72)],
                  npfOrig[, (83:88)], npfOrig[, (99:105)])

npfsp <- npf.168[, c(1:2, 5:6, 9:10, 13:16, 19:20, 23:24, 27:30, 37)]
colnames(npfsp)
npfsp.fact <- npfsp
npfsp.fact[, 19] <- as.factor(npfsp.fact[, 19])

npf4<-npfOrig[,-c(1,2,4)]
npf4.168 <- data.frame(npf4[, (1:3)], npf4[, (10:13)], npf4[, (24:27)], 
                       npf4[, (38:39)], npf4[, (50:51)], npf4[, (60:69)],
                       npf4[, (80:85)], npf4[, (96:101)])
npf4sp<-npf4.168[,c(1,2:3,6:7,10:11,14:17,20:21,24:25,28:31)]
```

```{r functions}
## Accuracy
accuracy <- function(pred, y) mean(ifelse(pred >= 0.5, 1, 0) == y)

## Perplexity
## Force probabilities to [0.995, 0.005]
perplexity <- function(pred, y) {
  pred <- ifelse(pred != 0 & pred != 1, pred, ifelse(pred == 0, 0.005, 0.995 ))
  exp(-mean(log(ifelse(y == 1, pred, 1 - pred)))) 
}

# Cross-validate
# Split n items into k folds of roughly equal size
kpart <- function(n, k) {
    rep_len(1:k, length.out = n)
}

# Cross-validation predictor (from week 1 problem 2)
cv <- function(
    formula, # variables to use
    data, # data
    model = lm, # model to train
    n = nrow(data), # rows in the matrix
    k = min(n, 10), # cross-validation folds
    split = kpart(n, k), # the split of n data items into k folds
    train = function(data) model(formula, data = data),
    pred = function(model, data) predict(model, newdata = data)) {
  yhat <- NULL
  for (i in 1:k) {
    mod <- train(data[split != i, ])
    if (is.null(yhat)) {
      yhat <- pred(mod, data)
    } else {
      yhat[split == i] <- pred(mod, data[split == i, ])
    }
  }
  return(yhat)
}

loocv <- function(
    formula, # Formula specifying which variables to use
    data, # Dataset
    model = lm, # Type of model to train (as a function)
    ## function to train a model on data
    train = function(data) model(formula, data = data),
    ## function to make predictions on the trained model
    pred = function(model, data) predict(model, newdata = data)) {
  yhat <- NULL
  for (i in 1:nrow(data)) {
    ## go through all stations, train on other stations, and make a prediction
    mod <- train(data[-i, ])
    if (is.null(yhat)) {
      ## initialise yhat to something of correct data type,
      yhat <- pred(mod, data)
    } else {
      yhat[i] <- pred(mod, data[i, ])
    }
  }
  yhat # finally, output cross-validation predictions
}

# Find cross-validation predictions, return probabilities and not the classification
cv.prob <- function(
               formula, # Formula specifying which variables to use
               data, # Dataset
               model = lm, # Type of model to train (as a function)
               n = nrow(data), # number of rows in the data matrix
               k = min(n, 10), # number of cross-validation folds
               split = kpart(n, k), # the split of n data items into k folds
               ## function to train a model on data
               train = function(data) model(formula, data = data),
               ## function to make predictions on the trained model
               pred = function(model, data) 
                 predict(model, newdata = data, type = "raw")[,2]) {
    yhat <- NULL
    for (i in 1:k) {
        ## go through all folds, train on other folds, and make a prediction
        mod <- train(data[split != i, ])
        if (is.null(yhat)) {
            ## initialise yhat to something of correct data type,
            yhat <- pred(mod, data)
        } else {
            yhat[split == i] <- pred(mod, data[split == i, ])
        }
    }
    yhat # finally, output cross-validation predictions
}

loocv.prob <- function(
               formula, # Formula specifying which variables to use
               data, # Dataset
               model = lm, # Type of model to train (as a function)
               ## function to train a model on data
               train = function(data) model(formula, data = data),
               ## function to make predictions on the trained model
               pred = function(model, data) 
                 predict(model, newdata = data, type = "raw")[,2]) {
    yhat <- NULL
    for (i in 1:nrow(data)) {
        ## go through all stations, train on other stations, and make a prediction
        mod <- train(data[-i, ])
        if (is.null(yhat)) {
            ## initialise yhat to something of correct data type,
            yhat <- pred(mod, data)
        } else {
            yhat[i] <- pred(mod, data[i, ])
        }
    }
    yhat # finally, output cross-validation predictions
}
```

## Multiple classes
### GLM
```{r glm, warning=FALSE, fig.height=4}
mod1 <- glm(class2 ~ ., npf, family = binomial(link = "logit"))
mod2 <- glm(class2 ~ ., npf, family = binomial(link = "probit"))
plot(mod1)
prob1 <- predict(mod1, npf, type = "response")
pred1 <- c()
perp1 <- c()
for (i in 1:dim(npf)[1]) {
  if (prob1[i] >= 0.5) {
    pred1 <- c(pred1, 1)
  } else {
    pred1 <- c(pred1, 0)
  }
  if (npf$class2[i] == 1) {
    perp1 <- c(perp1, prob1[i])
  } else {
    perp1 <- c(perp1, 1-prob1[i])
  }
}
```

GLM accuracy and perplexity. 
```{r glm res, warning=FALSE}
mean(pred1 == npf$class2)
exp(-mean(perp1))
```

### Lasso / Ridge
```{r lasso, message=FALSE, fig.height=4}
library(glmnet)
mod3 <- cv.glmnet(x = as.matrix(npf[, 1:100]), y = npf[, 101], alpha = 1, 
                  family = "binomial", type.measure = "class")
mod4 <- cv.glmnet(x = as.matrix(npf[, 1:100]), y = npf[, 101], alpha = 0, 
                  family = "binomial", type.measure = "class")
mod3a <- glmnet(x = as.matrix(npf[, 1:100]), y = npf[, 101], alpha = 1, 
                family = "binomial", lambda = mod3$lambda.min)
coef(mod3)
plot(mod3a)
plot(mod3$glmnet.fit, xvar = "lambda")
prob3a <- predict(mod3a, newx = as.matrix(npf[, 1:100]), type = "response")
pred3a <- c()
perp3a <- c()
for (i in 1:dim(npf)[1]) {
  if (prob3a[i] >= 0.5) {
    pred3a <- c(pred3a, 1)
  } else {
    pred3a <- c(pred3a, 0)
  }
  if (npf$class2[i] == 1) {
    perp3a <- c(perp3a, prob3a[i])
  } else {
    perp3a <- c(perp3a, 1-prob3a[i])
  }
}
```

Lasso accuracy and perplexity. 
```{r lasso  res, warning=FALSE}
mean(pred3a == npf$class2)
exp(-mean(perp3a))
```

```{r ridge, message=FALSE, fig.height=4}
mod4a <- glmnet(x = as.matrix(npf[, 1:100]), y = npf[, 101], alpha = 0, 
                family = "binomial", lambda = mod4$lambda.min)
plot(mod4a)
plot(mod4$glmnet.fit, xvar = "lambda")
prob4a <- predict(mod4a, newx = as.matrix(npf[, 1:100]), type = "response")
pred4a <- c()
perp4a <- c()
for (i in 1:dim(npf)[1]) {
  if (prob4a[i] >= 0.5) {
    pred4a <- c(pred4a, 1)
  } else {
    pred4a <- c(pred4a, 0)
  }
  if (npf$class2[i] == 1) {
    perp4a <- c(perp4a, prob4a[i])
  } else {
    perp4a <- c(perp4a, 1-prob4a[i])
  }
}
```

Ridge accuracy and perplexity. 
```{r ridge res, warning=FALSE}
mean(pred4a == npf$class2)
exp(-mean(perp4a))
```

### SVM
```{r svm, warning=FALSE}
modsv <- tune(svm, class2 ~ ., data = npf.fact, kernel = "linear", 
              ranges = list(cost = c(0.001, 0.01, 0.1, 1, 2, 5)))
modsvm <- svm(class2 ~ ., data = npf.fact, kernel = "linear", scale = FALSE, 
              cost = 0.001, probability = TRUE)
predsv <- predict(modsvm, newdata = npf.fact)
predsv2 <- predict(modsvm, newdata = npf.fact, probability = T)
probsv <- attr(predsv2, "probabilities")
perpsv <- c()
for (i in 1:dim(npf)[1]) {
  if (npf$class2[i] == 0) {
    perpsv <- c(perpsv, probsv[i, 2])
  } else {
    perpsv <- c(perpsv, probsv[i, 1])
  }
}
```
SVM accuracy and perplexity. 
```{r SVM res, warning=FALSE}
mean(predsv == npf$class2)
exp(-mean(log(perpsv)))
```

### Lasso (first version of final model)
```{r lasso2}
lasso1 <- cv.glmnet(x = as.matrix(npf[,1:100]), y = npf$class2, alpha = 1, 
                    family = "binomial", type.measure = "class", nfolds = 10)
lassomod1 <- glmnet(x = as.matrix(npf[,1:100]), y = npf$class2, alpha = 1, 
                    lambda = lasso1$lambda.min, family = "binomial", 
                    type.measure = "class")
preds <- c()
perps <- c()
for(i in 1:464) {
  newexp <- npf[-i, ]
  newmod <- glmnet(x = as.matrix(newexp[, 1:100]), y = newexp[, 101], 
                   alpha = 1, family = "binomial", lambda = lasso1$lambda.min)
  newprob <- predict(newmod, newx = as.matrix(npf[i, 1:100]), type = "response")
  if (npf$class2[i] == 1) {
    newperp = log(newprob)
  } else {
    newperp = log(1-newprob)
  }
  perps <- c(perps, newperp)
  if (newprob > 0.5) {
    newpred <- 1
  } else {
    newpred <- 0
  }
  preds <- c(preds, newpred == npfsp$class2[i])
}
mean(preds == npf$class2)
exp(-mean(perps))
```

### GLM (first version of actual model without usage of PCAs):
```{r glm2}
logreg <- glm(class2 ~ ., npfsp, family = binomial())
logreg <- glm(class2 ~ . -O3168.mean -O3168.std, npfsp, family = binomial())
logreg <- glm(class2 ~ . -O3168.mean -O3168.std -PTG.mean -PTG.std, npfsp, 
              family = binomial())
logreg <- glm(class2 ~ . -O3168.mean -O3168.std -PTG.mean -PTG.std -Pamb0.mean
              -H2O168.std -SO2168.mean-SO2168.std-1, npfsp, family=binomial())
preds <- c()
perps <- c()
for (i in 1:464) {
  newexp <- npfsp[-i,]
  newmod <- glm(class2 ~ . -O3168.mean -O3168.std -PTG.mean -PTG.std -Pamb0.mean 
                -H2O168.std -SO2168.mean -SO2168.std -1, data = npfsp, 
                family = binomial())
  newprob <- predict(newmod, newdata = npfsp[i, ], type = "response")
  if (npf$class2[i] == 1) {
    newperp = log(newprob)
  } else {
    newperp = log(1-newprob)
  }
  perps <- c(perps,newperp)
  if (newprob > 0.5) {
    newpred <- 1
  } else {
    newpred <- 0
  }
  preds <- c(preds, newpred == npfsp$class2[i])
}
mean(preds)
exp(-mean(perps))
```

### SVM (first version of actual model)
```{r svm2, eval=FALSE}
tc <- tune.control(cross = 464)
tuner <- tune(svm.class2 ~ ., data = npf.168.fact, kernel = "linear", type = "C",
              ranges = list(cost = seq(0.01, 0.05, 0.01)), tunecontrol = tc)
bpam <- tuner$best.parameters
preds <- c()
perps <- c()
for (i in 1:464) {
  newexp <- npf.168.fact[-i, ]
  newmod <- svm(class2 ~ ., newexp, kernel = "linear", cost = bpam[1, 1], 
                scale = F, probability = T, type = "C")
  newprob <- predict(newmod, newdata = npf.168.fact[i, 1:36], probability = T)
  if (class2[i] == 1) {
    newperp <- attr(newprob, "probabilities")[1, 1]
  } else {
    newperp <- attr(newprob, "probabilities")[1, 2]
  }
  perps <- c(perps, log(newperp))
  preds <- c(preds, newprob[1])
}
mean((preds-1) == class2)
exp(-mean(perps))
```

Logistinen regressio, 2 luokkaa
```{r}
logreg<-glm(class2~., npfsp, family=binomial())
logreg<-glm(class2~.-O3168.mean-O3168.std, npfsp, family=binomial())
logreg<-glm(class2~.-O3168.mean-O3168.std-PTG.mean-PTG.std, npfsp, family=binomial())
logreg<-glm(class2~.-O3168.mean-O3168.std-PTG.mean-PTG.std-Pamb0.mean-H2O168.std-SO2168.mean-SO2168.std-1, npfsp, family=binomial())
predsc2<-c()
perpsc2<-c()
for(i in 1:464){
  newexp<-npfsp[-i,]
  newmod<-glm(class2~.-O3168.mean-O3168.std-PTG.mean-PTG.std-Pamb0.mean-H2O168.std-SO2168.mean-SO2168.std-1, data=npfsp, family=binomial())
  newprob<-predict(newmod, newdata=npfsp[i,],type="response")
  if(npfsp$class2[i]==1){
    newperp=log(newprob)
  } else{
    newperp=log(1-newprob)
  }
  perpsc2<-c(perpsc2,newperp)
  if(newprob>0.5){
    newpred<-1
  } else{
    newpred<-0
  }
  predsc2<-c(predsc2, newpred)
}
mean(preds)
exp(-mean(perps))
```

Logistinen regressio, 4 luokkaa
```{r}
preds<-c()
perps<-c()
npf4sp$class4<-relevel(npf4sp$class4, ref="nonevent")
multmod<-multinom(class4~., npf4sp)
for(i in 1:464){
  newtrain<-npf4sp[-i,]
  newval<-npf4sp[i,]
  newmod<-invisible(multinom(class4~., newtrain))
  probs<-predict(newmod, newval, type="probs")
  if(probs[1]==max(probs)){
    preds<-c(preds,"nonevent")
  } else if (probs[2]==max(probs)){
    preds<-c(preds,"Ia")
  } else if (probs[3]==max(probs)){
    preds<-c(preds,"Ib")
  } else{
    preds<-c(preds,"II")
  }
  if(npf4sp$class4[i]=="nonevent"){
    perps<-c(perps, log(probs[1]))
  } else if (npf4sp$class4[i]=="Ia"){
    perps<-c(perps, log(probs[2]))
  } else if (npf4sp$class4[i]=="Ib"){
    perps<-c(perps, log(probs[3]))
  } else {
    perps<-c(perps, log(probs[4]))
  }
}
```